---
title: "Week 10- Resilience of tree growth"
output: html_document
date: "2026-02-08"
---

2.	Test the hypothesis that the resilience of tree growth (resilience.biomass.production) depends on species richness in Spanish plots. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
FunDivEurope<-read.csv("../Data/FunDivEUROPE.csv")
library(nlme)
library(emmeans)
library(ggplot2)
```

```{r}
#prepare data subset
spanish_plots <- subset(jena_annual, country == "Spain")

head(spanish_plots)
```
```{r}
# Basic LM
lm.resilience <- lm(resilience.biomass.production ~ target.species.richness, data = spanish_plots)
summary(lm.resilience)
```
Interpretation: Tree growth resilience increased significantly with increasing species richness in Spanish plots (linear model: p = 0.042).


```{r}
#check residuals
plot(lm.resilience)       
qqnorm(resid(lm.resilience))
qqline(resid(lm.resilience))

#homogenous
gls.resilience <- gls(resilience.biomass.production ~ target.species.richness,
                      data = spanish_plots,
                      weights = varIdent(form = ~1 | target.species.richness))
anova(gls.resilience)
```

When accounting for heteroscedasticity using a generalized least squares model, the effect of species richness on resilience was no longer statistically significant (p = 0.11).


